---
title: "Practical Machine Learning"
author: "PY Chia"
date: "Monday, February 23, 2015"
output: html_document
---

## Load Data , Clean Data and Partition

Training data is downloaded from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

Test data is loaded from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r, warning=FALSE}
library(caret)
set.seed(13333)
pmlTrain<-read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
pmlTest<-read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))
```

## Data set is cleaned and partitioned

Partitioning was performed to obtain a 75% training set and a 25% test set.

```{r,, warning=FALSE}
## Exclude NAs
noNApmlTrain<-pmlTrain[, apply(pmlTrain, 2, function(x) !any(is.na(x)))] 

## Exclude time and user information
pmlTrain1<-noNApmlTrain[,-c(1:8)]
pmlTest1<-pmlTest[,names(pmlTrain1[,-52])]

#data cleaning, Partitioning was performed to obtain a 75% training set and a 25% test set.

inTrain<-createDataPartition(y=pmlTrain1$classe, p=0.75,list=F)
training<-pmlTrain1[inTrain,] 
test<-pmlTrain1[-inTrain,] 

#Training and test set dimensions
dim(training)
dim(test)

```

## Using ML algorithms for prediction: Random Forest

Random forest trees were generated for the training dataset using cross-validation. The generated algorithm was later examined under the partitioned training set to check the accuracy and estimated error.

```{r,warning=FALSE}
fitControl2<-trainControl(method="cv", number=5, allowParallel=T, verbose=T)
rffit<-train(classe~.,data=training, method="rf", trControl=fitControl2, verbose=F)

predrf<-predict(rffit, newdata=test)
confusionMatrix(predrf, test$classe)

pred20<-predict(rffit, newdata=pmlTest1)
# Output for the prediction of the 20 cases provided
pred20

```

## Write to Output file

The results of the 20 predictions were then written to the output files.

```{r}


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(pred20)

```

